Run-Time Monitoring with Adjustable Overheads \\Using Dataflow-Guided Filtering

### 2014-12-18 dropping.tex
Update to camera-ready formatting.
Various edits throughout paper.

Figure~\ref{fig:dropping.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitor} performs a
series of \emph{monitoring operations}. The monitor operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of monitoring event, different monitoring operations are
executed. Information about monitoring events is sent to the monitor and
buffered in a FIFO structure to decouple the
running of the main core and the monitor. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitor may take several cycles to process a single event
from the main core. 
Monitors typically maintain metadata corresponding to each memory location
({\tt mem\_metadata[addr]}) and register ({\tt rf\_metadata[reg\_id]}) of the
main core.
If the monitor
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

@@

% Example monitor
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/example_full.pdf}
    \vspace{-0.2in}
    \caption{Example of array bounds check monitor.}
    \label{fig:dropping.example_full}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

Note that this monitoring is performed automatically and transparently with
almost no modification needed to the main program. The only modifications to
the main program that are needed are to initially set metadata, such as setting
the base and bounds addresses on a {\tt malloc} call for array bounds check.
Later propagation and checks of the metadata occur automatically as
instructions are forwarded.
Here, we have shown the dynamic sequence of operations executed by the monitor but
the actual static code consists of a set of instructions to be run for each
possible instruction type (load, store, etc.). Only instruction types relevant
for the particular monitoring scheme are forwarded.

@@

On the other hand, if the monitoring for accessing memory location {\tt y+12} on
line 4 is skipped, then
a false negative occurs. Originally, the monitor would catch this access as an
out-of-bounds access and raise an error. However, if the monitoring operation
for this is dropped, then the error is not detected. This is the trade-off that
we make in order to reduce overheads. That is, instead of either 100\% coverage
with all the associated overheads or no coverage and no overheads, we 
enable middle points of partial coverage with some fraction of the full
overheads.

@@

The key cause of false positives is dropping monitoring operations that update
metadata. Dropping monitoring operations that check for an error (such as the check
for line 4 in Figure~\ref{fig:dropping.example_full}) can only cause
false negatives but will never cause false positives. On the other hand,
skipping monitoring operations that update metadata can lead to false positives
and false negatives.
Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata. Thus, our general approach for preventing false
positives is to associate a 1-bit flag with each metadata in order to mark these
metadata as valid or
invalid. Furthermore, this invalidation information is propagated in the same
way that metadata is. Figure~\ref{fig:dropping.example_invalid}
shows an example of associating invalid flags with metadata. Suppose that the monitoring for
line 1 is dropped. When a monitoring event is dropped, metadata that would have
been updated is marked as invalid. In this case, instead of the normal
operation of marking {\tt rf\_invalid[1]} as {\tt false}, it is instead
marked as {\tt
true}. Thus, when line 3 is reached, the monitoring event is dropped
since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
performed on line 4 to also be dropped. This is necessary because an error
would have been raised even if the access on line 4 was in bounds.

@@

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/dataflow_architecture.pdf}
    \vspace{-0.3in}
    \caption{Hardware architecture of the dataflow engine.}
    \label{fig:dropping.dataflow} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

One way to reduce the number of monitoring events that must be handled by the
monitor is to filter out monitoring events that correspond to operating on null
metadata. Null metadata corresponds to events that
are not relevant to the monitor. For example, Hardbound
\cite{hardbound-asplos08} filters out operations on non-pointer (i.e., no base
and bounds metadata) instructions since it is not relevant to array bounds
checking. More recently, FADE \cite{fade-hpca14} has been proposed as a general
hardware module to perform this null metadata filtering for a variety of
monitoring schemes.

Figure~\ref{fig:dropping.example_null} shows an example of how this null filtering operates.
Here, the main core's code has been slightly modified from
Figure~\ref{fig:dropping.example_full} and {\tt r2} on line 2 is no longer set as an array pointer.
Without null metadata filtering, the instruction for line 4 would be
forwarded to the monitor since the system does not know whether {\tt r2}
contains a pointer address or not. However, if we use a 1-bit flag to mark {\tt r2}
as null when it is loaded with a constant, then we can propagate this null
information and filter out monitoring for line 4.

>>>>

Figure~\ref{fig:dropping.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitor} performs a
series of \emph{monitoring operations}. The monitor operates in parallel to the
main core. Events that trigger monitoring are referred to as \emph{monitoring events}. Depending
on the type of monitoring event, different monitoring operations are
executed. Information about monitoring events is sent to the monitor and
buffered in a FIFO structure to decouple the
running of the main core and the monitor. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. These stalls are a major source of
overhead because the monitor may take several cycles to process a single event
from the main core. 
Monitors typically maintain metadata corresponding to each memory location
({\tt mem\_metadata[addr]}) and register ({\tt rf\_metadata[reg\_id]}) of the
main core.
If the monitor
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

@@

% Example monitor
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/example_full.pdf}
    \caption{Example of array bounds check monitor.}
    \label{fig:dropping.example_full}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

Note that this monitoring is performed automatically and transparently with
almost no modification needed to the main program. The only modification to
the main program that is needed is to initially set metadata, such as setting
the base and bounds addresses on a {\tt malloc} call for array bounds check.
The propagation and checks of the metadata occur automatically as
instructions are forwarded.
Here, we have shown the dynamic sequence of operations executed by the monitor but
the actual static code consists of a set of instructions to be run for each
possible instruction type (load, store, etc.). Only instruction types relevant
for the particular monitoring scheme are forwarded.

@@

On the other hand, if the monitoring for accessing memory location {\tt y+12} on
line 4 is skipped, then
a false negative occurs. Originally, the monitor would catch this access as an
out-of-bounds access and raise an error. However, if the monitoring operation
for this is dropped, then the error is not detected. This reflects the trade-off that
we make in order to reduce overheads. Instead of either 100\% coverage
with all the associated overheads or no coverage and no overheads, we 
enable middle points of partial coverage with some fraction of the full
overheads.

@@

The key cause of false positives is dropping monitoring operations that update
metadata. Dropping monitoring operations that check for an error (such as the check
for line 4 in Figure~\ref{fig:dropping.example_full}) can only cause
false negatives and will never cause false positives. On the other hand,
skipping monitoring operations that update metadata can lead to false positives
and false negatives.
Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata. Thus, our general approach for preventing false
positives is to associate a 1-bit flag with each metadata in order to mark these
metadata as valid or
invalid. Furthermore, this invalidation information is propagated in the same
way that metadata is. Figure~\ref{fig:dropping.example_invalid}
shows an example of associating invalid flags with metadata. Suppose that the monitoring for
line 1 is dropped in order to meet the overhead target. When a monitoring event is dropped, metadata that would have
been updated is marked as invalid. In this case, instead of the normal
operation of marking {\tt rf\_invalid[1]} as {\tt false}, it is instead
marked as {\tt
true}. Thus, when line 3 is reached, the monitoring event is dropped
since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
performed on line 4 to also be dropped. This is necessary because an error
would have been raised even if the access on line 4 was in bounds.

@@

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/dataflow_architecture.pdf}
    \vspace{-0.1in}
    \caption{Hardware architecture of the dataflow engine.}
    \label{fig:dropping.dataflow} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

One way to reduce the number of monitoring events that must be handled by the
monitor is to filter out monitoring events that correspond to operating on null
metadata. Null metadata corresponds to events that
are not relevant to the monitor. For example, Hardbound
\cite{hardbound-asplos08} filters out operations on non-pointer (i.e., no base
and bounds metadata) instructions since it is not relevant to array bounds
checking. More recently, FADE \cite{fade-hpca14} has been proposed as a general
hardware module to perform this null metadata filtering for a variety of
monitoring schemes. Our architecture is able to support this null metadata
filtering with a small modification.

Figure~\ref{fig:dropping.example_null} shows an example of how this null filtering operates.
Here, the main core's code has been slightly modified from
Figure~\ref{fig:dropping.example_full} and on line 2, {\tt r2} is no longer set as an array pointer.
Without null metadata filtering, the instruction for line 4 would be
forwarded to the monitor since the system does not know whether {\tt r2}
contains a pointer address or not. However, if we use a 1-bit flag to mark {\tt r2}
as null when it is loaded with a constant, then we can propagate this null
information and filter out monitoring for line 4.

### 2014-12-18 evaluation.tex
Update to camera-ready formatting.
Various edits throughout paper.

We implemented our dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We implement the monitor as a software-based monitor
running on a parallel processor core, similar to LBA \cite{lba-asid06}.
We model the main and monitoring cores as
running at 2.5
GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for invalid and null flags.

In order to explore the generality of the architecture for
different monitors, we implemented five different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), instruction-mix profiling (IMP), and LockSet-based race detection (LS).  Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
described in Section~\ref{sec:dropping}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations or to detect malicious
software \cite{tang-raid14}.
LockSet \cite{eraser-tocs97} attempts to detect possible race conditions in multi-threaded programs
by tracking metadata about which locks are protecting shared memory
locations. If a shared memory location is accessed while unprotected
then a race condition may exist. Since we implement the monitor using a
processor core and our dataflow engine was designed to be generally applicable,
the same hardware platform supports all of these monitoring techniques.

@@

Since LockSet race detection requires multi-threaded programs, we tested it using
the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
200 million instructions eachi using 2 main cores to run the benchmark application.
Each main core was connected to a dataflow engine and a monitoring core.

@@

% Filter monitoring overheads
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0.1in]{figs/data_filtering.pdf}
    \vspace{-0.2in}
    \caption{Monitoring overheads with null metadata filtering.}
    \label{fig:evaluation.filtering}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.filtering} shows the
execution times of performing monitoring with null filtering enabled normalized
to the execution time of the benchmarks without monitoring. UMC sees normalized
execution times of 1.7-7x with an average of 4x.
For BC, normalized execution
are 3x on average and range from 1.1x to 8.3x.
% This is not surprising since in the baseline
% implementation all loads and stores needed to be monitored. However, with
% filtering, only loads and stores corresponding to arrays need to be forwarded.
DIFT sees an average of 1.1x slowdown with a maximum slowdown of 1.3x with null
metadata filtering. This low overhead is due, in part, to the fact
that for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would observe higher overheads. IMP shows normalized
execution times of 2-6x with an average of 4x. Overheads for LS (not pictured)
applied to the SPLASH-2 benchmarks are
1.03-1.72x after null metadata filtering and the average
overhead observed is 1.33x.
Our baseline system and overheads are similar to previous multi-core
monitoring systems such as LBA-accelerated \cite{lba-isca08} and FADE
\cite{fade-hpca14} (see
Table~\ref{tab:monitoring.previous_overheads}). In
Section~\ref{sec:evaluation.fpga}, we also evaluate a higher performance, FPGA-based
monitor that shows low tens of percent of overhead.

@@

% BC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for BC.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for UMC.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% LS sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_ls_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for LS.}
    \label{fig:evaluation.ls_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% IMP sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_imp_sweep.pdf}
    \vspace{-0.2in}
    \caption{Average error vs. varying overhead budget for instruction-mix profiling. Whisker bars show min and max errors.}
    \label{fig:evaluation.imp_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

In this section, we evaluate the effectiveness of
using partial monitoring to trade-off coverage for reduced overheads.
For these results, we use a slack-based, unrestricted dropping policy.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. 
We define \emph{monitoring coverage} as the 
percentage of dynamic checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC and LS). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system, 
assuming that errors/attacks are uniformly distributed across checks. 
%While we could not evaluate real errors/attacks due to difficulty in setting up
%a large number of bugs and exploits, we believe that the percentage of checks
%provides a good estimate of detection probability when errors/attacks are 
%uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.
In the figure, the bottom portion of each
bar shows the coverage for a 1.1x overhead target and the additional
coverage for increased overhead targets are stacked above this.

@@

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
average. Even higher coverage can be achieved by allowing higher overheads.
Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
With a 1.1x overhead, LS achieves 53\% coverage on average. This increases to
71\% coverage with a 1.2x overhead. 
We do not show a graph for DIFT because the overhead target of 1.1x is enough for
all benchmark tested to reach 100\% coverage.
%Since the overheads of DIFT are low to begin with, setting an overhead target
%1.1x is enough for all benchmarks tested to reach 100\% coverage.

@@

\begin{figure}
  \begin{center}
  \vspace{-0.1in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_exec_time.pdf}
    \label{fig:evaluation.umc_exec_time}
  }
  \vspace{-0.1in}
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_coverage.pdf}
    \label{fig:evaluation.umc_coverage}
  }
  \vspace{-0.1in}
  \caption{Comparing dropping policies for UMC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.umc_policies}
\end{figure}

In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
The results are shown for an overhead target of 1.5x.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overheads for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. We see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of short, independent monitoring chains.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
achieves higher coverage than unrestricted dropping. On average, unrestricted
dropping achieves 46\% coverage while source-only dropping achieves 49\%
coverage. This is due
to the fact that some of the overheads of monitoring for unrestricted dropping
are being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

@@

\begin{figure}
  \begin{center}
  \vspace{-0.1in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_exec_time.pdf}
    \label{fig:evaluation.bc_exec_time}
  }
  \vspace{-0.1in}
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_coverage.pdf}
    \label{fig:evaluation.bc_coverage}
  }
  \vspace{-0.1in}
  \caption{Comparing dropping policies for BC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.bc_policies}
\end{figure}

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_multirun_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
    \label{fig:evaluation.multirun}
    \vspace{-0.1in}
  \end{center}
\end{figure}

>>>>

We implemented our dataflow-guided monitoring architecture by
modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
run-time monitoring. We implement the monitor as a software-based monitor
running on a parallel processor core, similar to LBA \cite{lba-asid06}.
We model the main and monitoring cores as
running at 2.5
GHz with 4-way set-associative 32 kB private L1 I/D caches and a shared 8-way 2 MB L2
cache. This setup is similar to the Snapdragon 801 processor commonly found in
mobile systems. The dataflow engine uses a 1 kB cache for invalid and null flags.

In order to explore the generality of the architecture for
different monitors, we implemented five different monitors: uninitialized
memory check (UMC), array bounds check (BC), dynamic information flow
tracking (DIFT), instruction-mix profiling (IMP), and LockSet-based race
detection (LS).  
Since we implement the monitor using a
processor core and our dataflow engine was designed to be generally applicable,
the same hardware platform supports all of these monitoring techniques.
Uninitialized memory
check seeks to detect loading from
memory locations that are not initialized first. Array bounds check, as
described in Section~\ref{sec:dropping}, is a monitoring scheme that aims to
detect buffer overflows where memory accesses go beyond the boundaries of an
array. We modify the implementation of {\tt malloc} to set base and bound
metadata information. Dynamic information flow tracking is a security
monitoring scheme
which detects when information from untrusted sources is used to affect the
program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
files as untrusted. We implemented a multi-bit DIFT scheme which marks
untrusted data with a 32-bit metadata identifier so
that if an error is detected, it is possible to have information about where
the data originated from. Instruction-mix profiling counts the number of ALU,
load, store, and control instructions that are executed. This profiling
information can be useful for performing optimizations or to detect malicious
software \cite{tang-raid14}.
LockSet \cite{eraser-tocs97} attempts to detect possible race conditions in multi-threaded programs
by tracking metadata about which locks are protecting shared memory
locations. If a shared memory location is accessed while unprotected
then a race condition may exist.

@@

Since LockSet race detection requires multi-threaded programs, we tested it using
the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
200 million instructions each using 2 main cores to run the benchmark application.
Each main core was connected to a dataflow engine and a monitoring core.

@@

% Filter monitoring overheads
\begin{figure}
  \begin{center}
    %\includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0.1in]{figs/data_filtering.pdf}
    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
    \vspace{-0.2in}
    \caption{Monitoring overheads with null metadata filtering.}
    \label{fig:evaluation.filtering}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:evaluation.filtering} shows the
execution times of performing monitoring with null filtering enabled normalized
to the execution time of the benchmarks without monitoring. UMC sees normalized
execution times of 1.7-7x with an average of 4x.
For BC, normalized execution
are 3x on average and range from 1.1x to 8.3x.
% This is not surprising since in the baseline
% implementation all loads and stores needed to be monitored. However, with
% filtering, only loads and stores corresponding to arrays need to be forwarded.
DIFT sees an average of 1.1x slowdown with a maximum slowdown of 1.3x with null
metadata filtering. This low overhead is due to the fact
that for our implementation of DIFT on SPEC
benchmarks, we only mark data read from files as tainted. Instead, if we
targeted network or streaming applications, which have larger amounts of
untrusted input data, we would observe higher overheads. IMP shows normalized
execution times of 2-6x with an average of 4x. Overheads for LS (not pictured)
applied to the SPLASH-2 benchmarks are
1.03-1.72x after null metadata filtering and the average
overhead observed is 1.33x.
Our baseline system and overheads are similar to previous multi-core
monitoring systems such as LBA-accelerated \cite{lba-isca08} and FADE
\cite{fade-hpca14} (see
Table~\ref{tab:monitoring.previous_overheads}). In
Section~\ref{sec:evaluation.fpga}, we also evaluate a higher performance, FPGA-based
monitor that shows low tens of percent of overhead.

@@

% BC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_hb_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for BC.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for UMC.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% LS sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_ls_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for LS.}
    \label{fig:evaluation.ls_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% IMP sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true]{figs/data_imp_sweep.pdf}
    \vspace{-0.2in}
    \caption{Average error vs. varying overhead budget for instruction-mix profiling. Whisker bars show min and max errors.}
    \label{fig:evaluation.imp_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

In this section, we evaluate the effectiveness of
using partial monitoring to trade-off coverage for reduced overheads.
For these results, we use a slack-based, unrestricted dropping policy.
Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
array bounds check as we vary the overhead budget. 
We define \emph{monitoring coverage} as the 
percentage of dynamic checks that are performed 
(indirect jumps in DIFT, loads in UMC, and memory accesses in BC and LS). 
The metric is chosen to understand
how likely an error/attack instance is to be detected on an individual system, 
assuming that errors/attacks are uniformly distributed across checks. While
this is not necessarily true for actual errors/attacks, we believe this metric
provides a good initial estimate of the detection capabilities of the system.
%While we could not evaluate real errors/attacks due to difficulty in setting up
%a large number of bugs and exploits, we believe that the percentage of checks
%provides a good estimate of detection probability when errors/attacks are 
%uniformly distributed across checks.
Note that this metric is different from the
percentage of monitoring events that are not dropped, which includes
non-check instructions.
In the figure, the bottom portion of each
bar shows the coverage for a 1.1x overhead target and the additional
coverage for increased overhead targets are stacked above this.

@@

Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
we see that varying overhead budgets enables partial monitoring. We see that
with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
average. Even higher coverage can be achieved by allowing higher overheads.
Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
With a 1.1x overhead, LS achieves 53\% coverage on average. This increases to
71\% coverage with a 1.2x overhead. 
For DIFT (not pictured), an overhead target of 1.1x is enough for
all benchmarks tested to reach 100\% coverage.
%Since the overheads of DIFT are low to begin with, setting an overhead target
%1.1x is enough for all benchmarks tested to reach 100\% coverage.

@@

\begin{figure}
  \begin{center}
  \vspace{-0.1in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_exec_time.pdf}
    \label{fig:evaluation.umc_exec_time}
  }
  \\
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_coverage.pdf}
    \label{fig:evaluation.umc_coverage}
  }
  \caption{Comparing dropping policies for UMC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.umc_policies}
\end{figure}

In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
The results are shown for an overhead target of 1.5x.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overheads for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. We see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of short, independent monitoring dependence chains.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
achieves higher coverage than unrestricted dropping. On average, unrestricted
dropping achieves 46\% coverage while source-only dropping achieves 49\%
coverage. This is due
to the fact that some of the overheads of monitoring for unrestricted dropping
are being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

@@

\begin{figure}
  \begin{center}
  \vspace{-0.1in}
  \subfloat[Overhead budget error.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_exec_time.pdf}
    \label{fig:evaluation.bc_exec_time}
  }
  \\
  \subfloat[Coverage.]{
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_coverage.pdf}
    \label{fig:evaluation.bc_coverage}
  }
  \caption{Comparing dropping policies for BC.}
  \vspace{-0.2in}
  \end{center}
  \label{fig:evaluation.bc_policies}
\end{figure}

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_multirun_coverage.pdf}
    \vspace{-0.1in}
    \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
    \label{fig:evaluation.multirun}
    \vspace{-0.1in}
  \end{center}
\end{figure}
